### 2024-05-23 thesis/ms-fletcher-porter.tex
Integrate the JFPS work into the thesis

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                            %%
%% An example for writting your thesis using LaTeX                            %%
%% Original version by Luis Costa,  changes by Perttu Puska                   %%
%% Support for Swedish added 15092014                                         %%
%%                                                                            %%
%% This example consists of the files                                         %%
%%         thesistemplate.tex (versio 2.01)                                   %%
%%         opinnaytepohja.tex (versio 2.01) (for text in Finnish)             %%
%%         aaltothesis.cls (versio 2.01)                                      %%
%%         kuva1.eps                                                          %%
%%         kuva2.eps                                                          %%
%%         kuva1.pdf                                                          %%
%%         kuva2.pdf                                                          %%
%%                                                                            %%
%%                                                                            %%
%% Typeset either with                                                        %%
%% latex:                                                                     %%
%%             $ latex opinnaytepohja                                         %%
%%             $ latex opinnaytepohja                                         %%
%%                                                                            %%
%%   Result is the file opinnayte.dvi, which                                  %%
%%   is converted to ps format as follows:                                    %%
%%                                                                            %%
%%             $ dvips opinnaytepohja -o                                      %%
%%                                                                            %%
%%   and then to pdf as follows:                                              %%
%%                                                                            %%
%%             $ ps2pdf opinnaytepohja.ps                                     %%
%%                                                                            %%
%% Or                                                                         %%
%% pdflatex:                                                                  %%
%%             $ pdflatex opinnaytepohja                                      %%
%%             $ pdflatex opinnaytepohja                                      %%
%%                                                                            %%
%%   Result is the file opinnaytepohja.pdf                                    %%
%%                                                                            %%
%% Explanatory comments in this example begin with                            %%
%% the characters %%, and changes that the user can make                      %%
%% with the character %                                                       %%
%%                                                                            %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\documentclass[english,12pt,a4paper,pdftex,eng,utf8]{aaltothesis}

%% To the \documentclass above
%% specify your school: arts, biz, chem, elec, eng, sci
%% specify the character encoding scheme used by your editor: utf8, latin1

\usepackage{graphicx}

%% Use this if you write hard core mathematics, these are usually needed
\usepackage{amsfonts,amssymb,amsbsy}

\usepackage{color}

\definecolor{xkcd_blue}{RGB}{3,67,223} % https://xkcd.com/color/rgb/

%% Use this if you want to get links and nice output. Works well with pdflatex.
\usepackage{hyperref}
\hypersetup{
  pdfpagemode=UseNone,
  pdfstartview=FitH,
  colorlinks=true,
  urlcolor=xkcd_blue,
  linkcolor=black,
  citecolor=black,
  pdftitle=A Model for Selecting Software Tools for Fluid Power Systems,
  pdfauthor=Fletcher Porter,
  pdfkeywords={
    Mechatronics,Control Engineering,Software Tooling,Stewart Platform,Pneumatic actuator
  }
}

\usepackage[ddmmyyyy]{datetime}
\renewcommand{\dateseparator}{.}

\usepackage{siunitx}
\usepackage{gensymb}
\usepackage{svg}
\usepackage{lipsum}  % for placeholder text
\usepackage{subcaption}

%% All that is printed on paper starts here
\begin{document}

%% Change the school field to specify your school if the automatically
%% set name is wrong
% \university{aalto-yliopisto}
% \university{aalto University}
% \school{Sähkötekniikan korkeakoulu}
% \school{School of Electrical Engineering}

%% Only for B.Sc. thesis: Choose your degree programme.
%\degreeprogram{Electronics and electrical engineering}
%%

%% ONLY FOR M.Sc. AND LICENTIATE THESIS: Specify your department,
%% professorship and professorship code.
%%
\department{Department of Mechanical Engineering}
\professorship{Mechatronics}
%%

%% Valitse yksi näistä kolmesta
%%
%% Choose one of these:
%\univdegree{BSc}
\univdegree{MSc}
%\univdegree{Lic}

%% Your own name (should be self explanatory...)
\author{Fletcher Porter}

%% Your thesis title comes here and again before a possible abstract in
%% Finnish or Swedish . If the title is very long and latex does an
%% unsatisfactory job of breaking the lines, you will have to force a
%% linebreak with the \\ control character.
%% Do not hyphenate titles.
%%
\thesistitle{A Model for Selecting Software Tools for Fluid Power Systems}

\place{Espoo}

%% For B.Sc. thesis use the date when you present your thesis.
%%
%% Kandidaatintyön päivämäärä on sen esityspäivämäärä!
\date{\today}

%% B.Sc. or M.Sc. thesis supervisor
%% Note the "\" after the comma. This forces the following space to be
%% a normal interword space, not the space that starts a new sentence.
%% This is done because the fullstop isn't the end of the sentence that
%% should be followed by a slightly longer space but is to be followed
%% by a regular space.
%%
\supervisor{Prof.\ Jari Vepsäläinen}

%% B.Sc. or M.Sc. thesis advisors(s). You can give upto two advisors in
%% this template. Check with your supervisor how many official advisors
%% you can have.
%%
\advisor{D.Sc. (tech) Olof Colonius}

%% Aalto logo: syntax:
%% \uselogo{aaltoRed|aaltoBlue|aaltoYellow|aaltoGray|aaltoGrayScale}{?|!|''}
%%
%% Logo language is set to be the same as the document language.
%% Logon kieli on sama kuin dokumentin kieli
%%
\uselogo{aaltoRed}{''}

%% Create the coverpage
%%
\makecoverpage

%% Note that when writting your master's thesis in English, place
%% the English abstract first followed by the possible Finnish abstract

%% English abstract.
%% All the information required in the abstract (your name, thesis title, etc.)
%% is used as specified above.
%% Specify keywords
%%
%% Kaikki tiivistelmässä tarvittava tieto (nimesi, työnnimi, jne.) käytetään
%% niin kuin se on yllä määritelty.
%% Avainsanat
%%
\keywords{Mechatronics,Control Engineering,Software Tooling,Stewart Platform,Pneumatic actuator}
%% Abstract text
\begin{abstractpage}[english]
Fluid power systems can be enormously complex. The chiefest complexity is in their control components due to the deep knowledge required for effective development, the need to interface with myriad heterogeneous components, and the opacity of its operation. In the IT sector, a tech stack is a known discretization of function within a system, typically used to categorize what software tools do. This enables disparate parts of a system to be developed separately. Inspired by IT tech stacks, this paper presents a model, the mechatronic tech stack, to mitigate this complexity by siloing control system behavior into discrete functions and classifying what tools can be used to meet them. The model is demonstrated on three case studies of diverse application. The theory is promising, but there are challenges in defining the most useful abstractions and interfaces.
\end{abstractpage}

%% Force a new page so that the possible English abstract starts on a new page
%%
%% Pakotetaan uusi sivu varmuuden vuoksi, jotta
%% mahdollinen suomenkielinen ja englanninkielinen tiivistelmä
%% eivät tule vahingossakaan samalle sivulle
\newpage

%% Preface
%%
%% Esipuhe
\mysection{Preface}
%\mysection{Esipuhe}
I'd like to thank my teachers for making me wise, my family and friends for making me me, and the wind for guiding me in interesting directions. \\

\vspace{5cm}
Otaniemi, \today

\vspace{5mm}
{\hfill Fletcher Porter \hspace{1cm}}

%% Force new page after preface
%%
%% Pakotetaan varmuuden vuoksi esipuheen jälkeinen osa
%% alkamaan uudelta sivulta
\newpage

%% Table of contents.
\thesistableofcontents

%% Symbols and abbreviations
\mysection{Abbreviations}

\begin{tabular}{ll}
  IT & Information Technology \\
  HTML & Hyper Text Markup Language \\
  SQL & Structured Query Language \\
  CAN & Controller Area Network \\
  ISO & International Organization for Standardization \\
  OSI & Open Systems Interconnection \\
  HTTP & Hypertext Transfer Protocol \\
  DNS & Domain Name System \\
  FTP & File Transfer Protocol \\
  IP & Internet Protocol \\
  IO & Input/Output \\
  RAM & Random Access Memory \\
  CPU & Central Processing Unit \\
  API & Application Programming Interface \\
  I2C & Inter-Integrated Circuit \\
  USB & Universal Serial Bus \\
  PLC & Programmable Logic Controller \\
  DAQ & Data Acquisition System \\
  IEC & International Electrotechnical Commission \\
  EFPA & Extension type Flexible Pneumatic Actuator \\
  EEPROM & Electronically-Erasable Programmable Read-Only Memory \\
\end{tabular}

%% Tweaks the page numbering to meet the requirement of the thesis format:
%% Begin the pagenumbering in Arabian numerals (and leave the first page
%% of the text body empty, see \thispagestyle{empty} below).
%% Additionally, force the actual text to begin on a new page with the
%% \clearpage command.
%% \clearpage is similar to \newpage, but it also flushes the floats (figures
%% and tables).
%% There is no need to change these
%%
\cleardoublepage
\storeinipagenumber
\pagenumbering{arabic}
\setcounter{page}{1}

\section{Introduction}

%% Leave first page empty
\thispagestyle{empty}

In order to study mechatronics systems, it is necessary to design and develop physical research testbeds. This is especially the case in mechatronic systems with fluid components, as their high-fidelity modelling is challenging and computationally expensive. When doing research with hardware systems, before a single research question can be asked of the hardware, researchers must assemble mechanical components, hook up power electronics, and route hoses for the hydraulic and pneumatic systems. Then the most complex of them all, the control system, must be added. Its tangle of spindly wires reaches all throughout the machine, meeting at the computer, the single most complex component of any fluid power system. The requirements placed on it are immense. In high-power systems, safety invariants must be held up perfectly lest disaster strike, and close monitoring is necessary due to the low stiffness of fluid systems compared to mechanical ones.

This paper improves the process of commissioning control systems for fluid power systems by reasoning about how to abstract behavior and components in control systems into layers, synthesizing a novel {\it mechatronic tech stack\/} theory. It will demonstrate in three case studies how this theory can be applied in fluid power systems. The tools used in these cases will be categorized into layers of the mechatronic tech stack.  The analysis shows that, while different, these cases all fit into the same structure.

\clearpage

\section{Theory}

This section constructs the backbone of this paper, namely the novel theory of a mechatronic tech stack. First, there will be a review of tech stacks and {\it abstraction layers\/} as they've been applied in the IT sector. Then there will be a proposal of a mechatronic tech stack determined by a standards-based set of principles.

\subsection{Technology Stacks in the IT Sector}

The concept of a tech stack, alternately called a solution stack or a software stack, is to enumerate the tasks done by a kind of software application and to suggest tools for those tasks~\cite{PranamStack2017}. More specifically, it defines a set of abstraction layers; each layer describes a type of task and contains within it the set of software tools that complete that task. The intention is to take the difficult task of tool selection and provide a structure through which practitioners can think about the functions and tooling needs of their application, freeing mental energy to critically assess which tools would be best for their technical and personnel needs.

\subsubsection{The Web Stack}

A ubiquitous example of a tech stack in IT is the \textit{web stack}, used for developing websites and related pieces of software. The task that websites must perform, as suggested by this stack, is to present an interface to a user through a web browser (the \textit{front end}), to conduct some business logic like creating a customized page for the user (the \textit{back end}), to perform network communication (the \textit{server}), and to store and dispatch data (the \textit{database})~\cite{PranamStack2017}. This doesn't describe all websites---many don't have databases or any real back-end logic, but it describes at least a superset of the functionality of most websites.

Tools in each of these layers are myriad and to highlight that, proceeding in figure~\ref{fig:web_stack} is a list of a small sample. Front-end tools like jQuery~\cite{jQuery} have been proven from many years of use, React~\cite{React} is very powerful and can be used in non-web contexts, and even plain HTML without an additional framework can be good on account of its simplicity. Back-end tools can allow a developer to select their favorite programming languages like Ruby or Python, or one that is specialized for web back-ends like PHP, or even JavaScript like the front end uses. Servers are often integrated into the back-end framework, but if not, specialized servers like Apache~\cite{ApacheServer} or nginx~\cite{nginx} are popular. Databases are dominated by reliable and proven relational, SQL-based offerings like MySQL or PostgreSQL, but alternatives like NoSQL or Redis~\cite{redis} have become popular in the past years.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.7\textwidth]{assets/web_stack}
  \caption{The web stack is like four bins of tools, here illustrated as boxes with the logos of common tools within them.}\label{fig:web_stack}
\end{figure}

In discussing the web stack, it's also useful to think about a particular set of tools used for a single project, i.e.\ one tool from each layer. Alternately, this can be thought of as the project's cross section of the web stack, or the project's own stack. There are several of these cross sections that are widely used, such as LAMP (Linux, Apache, MySQL, PHP) or MEAN (MongoDB, Express.js, Angular.js, Node.js)~\cite{PranamStack2017}. Just as often, firms employ stacks tailored to their technical and personnel needs and often write about these decisions in blog posts, such as for recent changes to Microsoft Teams in~\cite{Singh2023}.

\subsubsection{OSI Network model}

Another example of a tech stack is the OSI Network model which abstractly describes network communication down to the lowest levels of physical signaling up to the highest levels of abstract behavior~\cite{ISO7498-1}. This model is frequently cited in mechatronic literature since many fieldbus protocols like CAN define their functionality in terms of this model. This model is also interesting since its standard description (ISO~7498-1) also includes principles for creating layered models of a similar ilk.

The OSI model differs from the web stack in how the layers are used by practitioners. Typically, instead of each layer being a bin to pick a tool from and where a tool must be picked from every layer which you want functionality, developers instead may decide at what depth into the model they begin their implementation and assume everything below that depth ``just works''.

The model itself consists of layered abstractions over physical communication media, as seen in figure~\ref{fig:osi_model}. The lowest layer, Physical, concerns itself with the physical phenomena of communication. In the case of CAN, this means voltage levels, terminating resistors, twisted pairs, grounding rules, and more. The highest level, Application, provides network facilities without need for understanding \textit{how} the network connection is made. The Application layer includes, for example, identifying communication partners and negotiating security protocols. Well-known protocols like HTTP, DNS, and FTP exist in the Application layer. The intermediate layers provide increasing levels of abstraction over the physical layer, defining things like the structure of data packets, how to route messages, how to route them efficiently, and more.

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{assets/osi_model}
  \caption{The OSI model understands network communication to be a layer cake of abstractions over the physical communication media.  Taken from~\cite[§6.1.3]{ISO7498-1}.}\label{fig:osi_model}
\end{figure}

An interesting property of layered abstraction architectures is that protocols which are defined in higher layers can have the lower layers transparently replaced. For instance, the internet protocol which we use every day is defined in the Network and Transport layers. Typically, the internet is built over Ethernet or Wi-Fi, but with some effort, other low-level protocols are available. Lindgren et al.\ developed IP over CAN, wherein an internet network is extended over a CAN network~\cite{Lindgren2008}, which was never imagined by implementers of CAN nor IP.

\subsubsection{Principles for Identifying Abstraction Layers}

The OSI standard also includes the principles upon which its layers were designed~\cite[§6.2.1]{ISO7498-1}. In summary, each layer should be a discrete, separable item with a coherent function. Their interfaces should be succinct and only connect with adjacent layers. Fewer layers eases explanation and implementation and the choice layers should be motivated by experience. Being a novel theory, experience will be the principal gain of future work, although case studies discussed later will also help. Usefully, it offers a test to aid in determining layer boundaries. Layers should have a different level of abstraction in terms of \textit{morphology}, \textit{syntax}, and \textit{semantics}.

This model of morphology, syntax, and semantics comes from Bloomfieldian linguistics~\cite{Bloomfield1923}. It's used to coerce natural language into a structured form from which meaning can be gleaned. Here, the standard authors are using it to glean structure and meaning from engineering systems. To understand this theory, take this line of C code.

\begin{verbatim}
printf("I have %u cows\n", num_cows);
\end{verbatim}

Morphology is concerned with the atomic elements of meaning. In C code, this means short tokens of text, here \verb|printf|, \verb|(|, \verb|"I have %u cows\n"|, \verb|,|, \verb|num_cows|, \verb|)|, and \verb|;|. None of these can be reduced and still carry meaning---\verb|pri| and \verb|ntf| doesn't make \verb|printf|. But these atoms of meaning needn't be text or language. As will be discussed, they can be cables, connectors, computer hardware, or anything with relational meaning.

When these tokens are combined into phrases, that's syntax. Syntax is the structure into which meaning is applied. The \verb|printf| example is syntactically parsed by C compilers as in figure~\ref{fig:printf_syntax}. The syntax tells that the postfix expression represented by the identifier printf has the argument list of \verb|"I have %u cows\n"| (a string literal) and \verb|num_cows| (an identifier). What any of these tokens mean is the domain of semantics.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.8\textwidth]{assets/printf_syntax}
  \caption{The above printf example with syntax labeled according to K{\&}R C~\cite{Kernighan1978}}\label{fig:printf_syntax}
\end{figure}

To understand the semantics of \verb|printf|, you must read the documentation. It will tell you that \verb|printf| means to convert and write output to the console under the control of a format string. You'll also see that \verb|"I have %u cows\n"| is that format string, and \verb|num_cows| is an additional argument, presumably a number of cows. Reading the documentation further, you'll see that the format string itself has its own morphology, syntax, and semantics driving how output is created.

Technology stacks, like the web stack or the OSI stack, create a structure for selecting tools for an application and abstractions for interfacing with other layers of the stack. The OSI model also provides guidance on how to architect a layered model based, among other things, on creating layers with coherent morphology, syntax, and semantics.

Next, these ideas will be used to make a mechatronic tech stack like the web and OSI stacks.

\subsection{A Mechatronic Tech Stack}

Broadly speaking, mechatronic systems include mechanical systems, power-electronic systems, and digital control systems. All of these could be considered under a theory of a mechatronic tech stack, but the present treatment will only consider the control systems. Specifically, the stack will start at the IO (input/output) interface for sensors and transducers.

The layers will be identified by noting contrasts in morphology, syntax, and semantics throughout the gamut of the control system. Layer boundaries will become evident where there are dramatic differences. I'll then describe the layers and the interfaces between them.

Morphology provides the strongest and most coarse distinctions in layers. If the atoms that one component is made of are incompatible with another, they can't effectively be considered together. The most obvious case is that those components that are made of literal atoms, i.e.\ hardware, must be considered separately from those that aren't physical, namely software. Within hardware, a control engineer will concern themselves with the components of the computer which their control software run on (RAM, CPU, storage, etc.) and the physical connections between the computer components and the sensors and transducers, such as cables, connectors, contacts, and traces. The software is the instructions of the control program itself, but also supporting software like the operating system, runtime environments, and libraries.

Syntax is the structure of an idea. It is a finer comb to discriminate layers. Software generally is built as a linear sequence of instructions with branches and loops. This describes the control engineer's perspective on their own control program, but supporting software like libraries are more abstract. They're better thought of as a dependency. A control program may depend on libraries A, B, and C, and those in turn depend on libraries D, E, F, and G. These \textit{middlewares} are better understood as a network of the programs themselves and the dependency connections between them. For hardware, its structure is its assemblage. The computer components are connected to the motherboard, sensors and transducers are connected to a CAN bus, the temperature sensors message over I2C, etc. Like middleware, hardware is the network of components and the connections between them, but the type of connection is also important. A system with components on a CAN bus is going to be different than one with a USB bus, and different still from a system where all components communicate over a serial connection.

Semantic meaning reveals the most subtle distinctions in layers. The control program is an expression of the business logic desired by the engineer. Middleware begins where the controller ends. It attempts to save the engineer from difficult, tedious, or uninteresting implementation otherwise necessary with computers. Computers are meaningful when they provide compelling output based on some input, i.e.\ when they compute. How this computation occurs at the hardware level is determined by the computer's \textit{architecture}. This defines what machine instructions do what and which hardware pins go where. IO is the hardware beyond the computer. It works on data transfer. A sensor's pressure reading becomes a message to the computer. The computer's actuation command becomes a message to the transducer. This is handled by e.g.\ network communication protocols, the domain of the OSI model.

This is enough to propose a mechatronic tech stack. The mechatronic stack as outlined sits on top of signaling with sensors and transducers. Those are interfaced by IO systems that connect to computer systems which run middleware upon which a control program depends. This is illustrated as a stack diagram in figure~\ref{fig:mechatronic_tech_stack}.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.7\textwidth]{assets/mechatronic_tech_stack}
  \caption{The mechatronic tech stack places high-level business logic at the top and low-level interfacing at the bottom.}\label{fig:mechatronic_tech_stack}
\end{figure}

Between the layers are the interfaces that enable higher layers to take advantage of the abstractions provided by lower layers. The control layer, being the highest, provides no interface to further layers, but from the middleware layer it demands certain \textit{application programming interfaces}, or APIs. Each program in the middleware layer generally provides an API, so the control-middleware interface is the union of all these APIs. The needs of middleware are typically met by other middleware, but those that interface with the computer directly do so through architecture-defined machine code. Then for the Computer-IO interface, the IO needs hardware connections like pins broken out from the CPU or connection ports like for CAN, serial, or Ethernet. Finally, IO interfaces with sensors and transducers in the same way it does with computers, through hardware interfaces like terminal blocks or hookups to a fieldbus.

This model, like the web stack, provides a structure through which to think about tool selection. The choice of using CAN- or serial-based IO can be understood separately from the decision of what programming language to use for the controller. This isn't to say that tooling decision in the high layers is completely decoupled from those in the low layers. The decision to use an AVR-based Arduino computer likely precludes writing the controller in JavaScript, but this is due to a lack of middleware to support that use case.

Following will be three case studies providing concrete illustrations of how this mechatronic tech stack model can be applied. These will mostly be descriptive case studies that apply this theory for retrospective research, but one case uses this model prescriptively.

\clearpage

\section{Case Studies}\label{sec:case_studies}

These case studies will apply the theory of a mechatronic tech stack to three projects---one where a requirement change caused a re-architecture of the control system, another where three similar devices have their stacks compared, and the last has its stack designed with this theory in mind. In all cases, the project's stack will be explicated and there will be a discussion of interesting features of its stack.

\subsection{Matlab or a PLC for a Steering Linkage?}

Dolores is a full-scale testbed in Aalto University's Fluid Power Laboratory, seen in figure~\ref{fig:dolores0} to study the application of electric hydrostatic actuation in an articulated-steering linkage for a front-loading mining vehicle~\cite{Hermansson2021, Helduser1999, Dudzinski1989}. Its control system commissioning was completed in 2021, but come 2023, an industry partner in the experiment requested hardware changes that demanded a substantial re-architecture of the control system. This case study will discuss the mechatronic tech stacks behind the old and the new systems.

\begin{figure}[h]
  \centering
  \includegraphics[width=0.6\textwidth]{assets/dolores}
  \caption{Dolores rotates the top platform about a pivot by actuating a pair of hydraulic cylinders.  A pair of additional cylinders in parallel can resist the motion providing loads useful for testing.}\label{fig:dolores0}
\end{figure}

The industry partner's requested change occurred in the IO (input/output) layer. The old system had an inverter that drives a servomotor, a safety valve to enable the system, and an array of pressure, temperature, and flow sensors to observe the state of the hydraulic system. The inverter and the pressure and temperature sensors all communicated over the CANopen protocol, the flow sensors gave analog inputs, and the safety valve was actuated with a digital signal. The CAN communication was interfaced by a Vector CAN/USB dongle~\cite{VectorCanUsbDongle}. All other IO went through a National Instruments data acquisition device (DAQ). In the new system, the forced change for the inverter to communicate over EtherCAT. This meant new IO hardware was needed to support this protocol, which was selected to be Beckhoff EtherCAT terminals. They provided the interface to connect the inverter, digital and analog IO, and a CAN bus for the sensor network~\cite{BeckhoffEL6751}. This made the DAQ unnecessary, but the CAN/USB dongle was still required due to deficiencies in Beckhoff's control interfaces, which are discussed with the control layer.

The system had been controlled by a conventional Windows workstation. This may have been fine to keep for the new system, but there was concern with using a whole workstation for a single experiment. Instead, one of Beckhoff's industrial PCs was selected as the computer to leverage Beckhoff's support~\cite{BeckhoffC60XX}. It was expensive, but it was thought that architecting the system to align with vendor expectations would make development easier and mitigate potential problems. However, the workstation did tap into the CAN bus to initialize the sensors, so it wasn't completely out of the loop.

In the old system, the workstation ran Windows, the DAQ was managed by drivers from National Instruments, and the CAN bus was initialized using Vector's CANalyzer application and runtime commands came through Vector's Simulink library~\cite{VectorCanalyzer, SimulinkVectorCAN}. Due to issues getting reliable EtherCAT connections with Simulink (possibly for user error, though the documentation wasn't simple), the new system uses a PLC-based middleware. Still based on Windows, it has libraries and runtimes to run the EtherCAT bus, to read from the CAN bus, and to interface with an external development computer. CANalyzer remained to initialize the CAN bus.

The old system had controllers written with Matlab and Simulink. However, due to middleware issues, they were abandoned in favor of writing controllers in PLC languages based on IEC~61131-3 with the TwinCAT development environment~\cite{IEC61131-3, BeckhoffTwinCAT}. The major issue with TwinCAT is that its CANopen implementation had no documented way to initialize devices. This was worked around by initializing devices with CANalyzer on the workstation.

Together, the stack for the old system was CANopen-based IO with a DAQ supplementing simple analog and digital signals, a workstation computer, Windows middleware with drivers and libraries to support communication with the CAN bus and DAQ, and a controller based on Matlab and Simulink. In the new system, the stack was a mix of EtherCAT and CANopen IO, a Beckhoff Industrial PC in concert with a workstation, PLC middleware with an application to initialize the CAN network, and a controller written in PLC languages with TwinCAT. This is illustrated in figure~\ref{fig:mechatronic_tech_stack_dolores}.

\begin{figure}[t]
  \centering
  \includegraphics[width=\textwidth]{assets/mechatronic_tech_stack_dolores}
  \caption{The change was to use Beckhoff's stack, though it didn't work without supplements.}\label{fig:mechatronic_tech_stack_dolores}
\end{figure}

Notably, a small change in the system, a change in IO protocol, demanded a wholesale re-architecture of the control system. Also notable is that issues in controller-layer tooling echoed through the whole stack. This speaks to how in both stacks, the tools are \textit{tightly coupled}, meaning changes in one tool often demand changes in the others. For this mechatronic stack model to be a useful means of selecting tools like the web stack, the layers need to be \textit{loosely} coupled so tools can be replaced with minimal change in other layers.

\subsection{Embedded Systems for Soft, Pneumatic Actuators}

A research group at Okayama University of Science has, over the past years, developed a novel soft, pneumatic actuator that extends when supplied with pressure called an EFPA~\cite{Noritsugu2005}. This section will study the tech stacks for three applications of this actuator, a six-legged mobile robot for providing the elderly with an interactive physical activity~\cite{Hase2020}, a pipe-traversal robot for inspecting and repairing water pipe networks~\cite{Shinohara2020}, and a lightweight servo valve based on bending hoses to support EFPA applications~\cite{Kobayashi2020}. All of these can be seen in figure~\ref{fig:soft_pneumatic_actuators_demo}. Note that because this is a case study on work we did not participate in, some details must be inferred.

\begin{figure}[h]
  \centering
  \begin{subfigure}[t]{0.31\textwidth}
    \centering
    \includegraphics[width=\textwidth]{assets/six_legged_robot}
    \caption{Six-legged robot~\cite{Hase2020}}\label{sfig:six_legged_robot}
  \end{subfigure}
  \begin{subfigure}[t]{0.35\textwidth}
    \centering
    \includegraphics[width=\textwidth]{assets/pipe_robot}
    \caption{Pipe-traversal robot~\cite{Shinohara2020}}\label{sfig:pipe_robot}
  \end{subfigure}
  \begin{subfigure}[t]{0.32\textwidth}
    \centering
    \includegraphics[width=\textwidth]{assets/bending_servo_valve}
    \caption{Bending servo valve~\cite{Kobayashi2020}}\label{sfig:bending_servo_valve}
  \end{subfigure}
  \caption{These projects are made by the same people with the same technologies using similar tools.}\label{fig:soft_pneumatic_actuators_demo}
\end{figure}

In the IO layer, all three are similar. The six-legged and pipe robots have an array of on/off valves commanded by digital signals to pressurize their EFPAs. The servo valve has its position set with a small motor, which is controlled by a PWM signal. These signals are propagated through direct connection to pins on the computer, although Shinohara et al.\ note that there are circuits between stepping up power. The robots received command signals from a workstation over a serial connection. Hase et al.\ note that there was an intermediate chip on the six-legged robot to connect over USB. The servo valve computer had no reported inputs, though several sensors were connected to a data recording device.

The computer used for the two robots was an SH7125, a 32-bit SuperH-based system on a chip, a device that has all necessary components for computation like CPU, RAM, and peripherals on a single die~\cite{SH7125}. The servo valve used an H8/3264, a 32-bit H8/300-based system on a chip~\cite{H83264}. These architectures, while unfamiliar in name, are common for applications like printer or hard drive controllers where cost and hardware IO are important. Both CPUs have inbuilt peripherals for IO where when the CPU writes a number to a register, the IO pins will change state to match its binary representation.

No middleware is mentioned in these papers. While it's in principle possible that they ran some embedded operating system, it's doubtful given all their controllers seem fairly simple, as discussed below. It is possible that they used some external libraries to aid in programming. A programming environment without an operating system is referred to as \textit{bare metal} in industry parlance because that's all there is to the middleware interface.

The language the controllers were written in was not mentioned in the papers, but it's likely they used an assembly language or C since major compilers support both of these architectures~\cite{GccSupportedArchitectures}. Hase et al.\ state that the control program they wrote was to cycle through a fixed sequence of states for the valves, perhaps switching sequences interactively over the serial connection.

Each of these three projects implemented a largely similar stack---IO based on direct connections to the CPU with supporting circuitry, an embedded system-on-a-chip computer, no middleware, and controllers written in C or assembly. By using a common stack, these researchers were able to skip over the tedious and difficult work of selecting tools and just use what is familiar to their institution and focus their energy on research.

\subsection{A Novel Stack for a Pneumatic Stewart Platform}

With this new theory of a mechatronic tech stack, it made sense to start a project to study what tools this model lends itself to when applied prescriptively. The project was a driving simulator chair that had been an unfinished student project a year prior~\cite{Bjoerklund2023}. The mechanical systems had been finished including all the sensors and transducers, so all that was needed was a control system.

\begin{figure}[h]
  \centering
  \begin{subfigure}[t]{0.43\textwidth}
    \includegraphics[width=\textwidth]{assets/rocking_chair}
  \end{subfigure}
  \quad
  \begin{subfigure}[t]{0.53\textwidth}
    \includegraphics[width=\textwidth]{assets/pneumatic_cylinder_test}
  \end{subfigure}
  \caption{The whole system, left~\cite{Bjoerklund2023}, is a chair on a Stewart platform made with pneumatic cylinders.  On the right is a cylinder driven under the present tech stack.}\label{fig:stewart_platform_demo}
\end{figure}

The system is a chair atop a Stewart platform whose linear actuators are single-acting pneumatic-muscle cylinders, seen in figure~\ref{fig:stewart_platform_demo}~\cite{Stewart1965, Caldwell1995}. The valve manifolds that were already made for this had one valve to connect the pneumatic muscle to an external pressure source and another to vent the muscle to atmosphere. When both valves are closed, it brakes the cylinder. There are six instances of this circuit, connected in series to a common pressure source. Each cylinder also has a position sensor exposed as a voltage signal from a potentiometer.

Tool selection started from the IO layer. The need was to switch 12 digital outputs for the valves and six analog inputs ($\qtyrange[range-units=single,range-phrase=..]{0}{24}{\volt}$) from the position sensors. It was also desirable for the work to be done quickly to have a demonstration in time for publication. This was met with EtherCAT IO cards from Beckhoff~GmbH, in large part thanks to their availability within our lab. They supply IO cards that mostly met the requirements; the digital outputs needed to be supplemented with flyback diodes for the inductive load and the analog inputs needed signals stepped down to $\qtyrange[range-units=single,range-phrase=..]{0}{10}{\volt}$ with resistors~\cite{BeckhoffEL2042,BeckhoffEL3062}. This meant routing signal cables from the sensors and transducers to terminal blocks, placing circuitry in the terminals, and then routing to the Beckhoff terminals, where there is an EtherCAT connection to the computer's Ethernet port~\cite{BeckhoffEK1100}.

In selecting the computer, it needed to communicate with the IO over EtherCAT and be cheap and easy to develop for. The Raspberry Pi supports EtherCAT, is reasonably priced, and Linux-based systems are familiar to the authors and easy to develop software for.

The principal middleware requirement was an EtherCAT master, which itself depends on system-level networking. Linux satisfies the networking requirement. EtherCAT masters are conventionally provided by PLC-based systems, like those from Beckhoff~\cite{BeckhoffTwinCAT}. However, they generally do not run on Linux. However, the point of this project was to try new things. So, Ethercrab, a master written in the Rust programming language, was used~\cite{Ethercrab}. Ethercrab benefits from being open source, so motivated users can troubleshoot more easily. It also has thorough documentation and lots of examples (e.g.\ the control-layer application was based on \verb|ethercrab/examples/ek1100.rs|). Interfacing with Rust is a benefit due to its ease of use, quality tooling, performance, and memory safety~\cite{RustHomePage}.

Finally, the control layer had to use the API of the middleware layer to implement the business logic. One case of that logic was to have a controller to regulate the position of each cylinder. It had to be implemented in Rust to interface with the Ethercrab library. It also required some code to initialize the EtherCAT communication and to configure the Beckhoff terminals. There's also a user interface to interactively set the cylinder position. An interesting issue came up where the position sensor signal, mapping the $\qtyrange[range-units=single,range-phrase=..]{0}{10}{\volt}$ IO to the range 0..1~\verb|float|, but only reached the range 0..0.5~\verb|float|. This was because the step-down resistor was larger than designed since the right size wasn't available, so the maximum sensor reading was smaller than $\qty{10}{\volt}$. This unhandled gain is an undesired interface between the Control and IO layers since it demands the controller be written with knowledge of the minutia of how the IO was implemented. Further research will go toward how to prevent this.

The stack used for this pneumatic Stewart platform was Beckhoff EtherCAT Terminals with supplemental circuits for IO, a Raspberry Pi computer, Linux/Ethercrab middleware, and the control layer was built with Rust.

\clearpage

\section{Discussion}

The mechatronic tech stack theory basically held to scrutiny through the case studies.  All cases presented all layers of the stack except the EFPA robots, which had no middleware layer, though that is an interesting insight. With this model, the commissioning issues with the steering linkage Dolores could partly be diagnosed as too-tight coupling between layers.  Commissioning the Stewart platform was basically easy; by separating functions into different categories, work on any layer could be done independently.  It also raises compelling questions; can Dolores' Beckhoff industrial computer be replaced with a Raspberry Pi (only if it would support Beckhoff's middleware)?  Could the controller for the Stewart platform have been written in C or a PLC language?  What would it take to change the fieldbus protocol without the hassle seen with Dolores?  These questions are natural to ask when the tooling is presented in a structured manner as with this model.

The main issue that arose was the IO layer reaching into higher layers.  Dolores' old control system failed because it couldn't support a change in communication protocol.  The new system is over-complex because the control tools struggle to support all communication protocols.  The Stewart platform also saw issues with this; measurement signals were unexpectedly low due to an off-size resistor in the IO layer which impacted control function.  Certainly, some knowledge of the details of the IO layer's implementation must be communicated to the middleware or control layers such that correct middleware is run and sensor calibrations are accounted for.  How this could be implemented will be studied in future work.

An additional issue is that the computer layer isn't overly interesting.  Each of the case studies simply used whatever computers are available or easy to use.  In some circumstances, it could be more interesting, like to discuss which computer architecture to use for an embedded application, or whether to use an embedded Linux system like a Raspberry Pi or bare metal, or even a simple EEPROM.  Perhaps the computer layer could steal operating systems away from the middleware layer, or these case studies weren't interesting for that layer.

\clearpage

\section{Conclusion}

This paper presented three case studies to demonstrate the utility of the novel theory of a mechatronic tech stack. These cases were diverse in their applications and methods and by viewing them through the lens of the tech stack, new knowledge of how fluid power systems can be commissioned was gained. Also, by studying several stacks, we have formed a pool of tools to draw from. The mechatronic tech stack can be re-rendered to reflect this, as here in figure~\ref{fig:mechatronic_tech_stack_filled}.

\begin{figure}[h]
  \centering
  \includegraphics[width=\textwidth]{assets/mechatronic_tech_stack_filled}
  \caption{Not only does the mechatronic tech stack present layers of abstraction, it also provides an idea of what tools to choose, similar to figure~\ref{fig:web_stack}.}\label{fig:mechatronic_tech_stack_filled}
\end{figure}

This model can be improved by conducting more retrospective analyses, like those done for Dolores and the EFPA projects, to better model existing systems. Further, executing more projects using this theory will test and exercise its utility. The scope of the model can also be increased; instead of only modeling control systems, it could be expanded to model mechanical, electrical, or fluid systems.

Using this model, practitioners can more easily reason about what different parts of their control systems do and what tools to employ for them. This mental model enables faster development of systems and redirects thought to what tools would be best instead of merely functional. All this will allay the difficulty and complexity of developing fluid power systems.

\clearpage

\phantomsection
\addcontentsline{toc}{section}{\refname}
%%\addcontentsline{toc}{section}{References}

\bibliographystyle{ieeetr}
\bibliography{bib/good-masters.bib}

\end{document}